The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-04-16 paper/iphas-dr2.tex
First attempt to incluto include Vega fluxes, data access section, and detection statistics paragraph properly.

Unlike the broadbands, 
standard-star photometry is not available in the literature 
for the H$\alpha$ passband
and hence there is no formally recognised flux scale 
for it.
We can specify here, however, 
that the integrated in-band energy flux for Vega 
in the IPHAS \ha\ filter 
is $1.52 \times 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$ 
at the top of the Earth's atmosphere,
which is the flux obtained by folding 
Vega's SED with the filter transmission curve 
(uncorrected for the atmosphere and detector quantum efficiency,
which would otherwise scale down the narrowband flux by 0.707).
This is 3.14 magnitudes less than the flux captured 
by the much broader $r$ band
which includes the \ha\ band.
Hence to assure zero colour relative to the broadbands,
we set the default zeropoint for the narrowband to be:
\begin{equation}
\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.14.
\label{eqn:zpha}
\end{equation}
Note that the above implies that the in-band flux corresponding to
zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
when the \ha\ magnitude for Vega is set by convention to 0.03.

@@

A total of 219 million primary detections satisfied
the above criteria and have been included in the catalogue.
Amongst these, 159 million are detected 
in all three bands (73 per cent),
30 million are detected in two bands (14 per cent),
and the remaining 30 million are single-band detections.
The detailed breakdown is presented in Table~\ref{tbl:detections},
along with the fraction of `confirmed' sources,
hich we define
as objects which were detected in more than
one field pointing (usually the alternative pointing of a field pair).
Sources detected in all three 
(Table~\ref{tbl:detections}).
Half of the single-band detections were made in the $i$-band.
This is most likely because $i$ is least
affected by interstellar extinction and can occasionally pick up
highly-reddened objects which are otherwise lost in $r$ and \ha.
Table~\ref{tbl:detections} also shows the fraction of the
sources that were `confirmed', w
For the class of 
in more than one field (usually the field observed at the small offset position)

\begin{table}
    \caption{Breakdown of catalogue entries as a function of the band(s) in which the source was detected. We also show the fraction of `confirmed' objects, which we define as sources which were detected in more than one field pointing (usually the field pair partner). Also included are the mean magnitudes for each class.}
    \label{tbl:detections}
        \begin{tabular}{lrcccc}
        \toprule
        Band(s) & Sources & Confirmed & $\overline{r}$ & $\overline{i}$ & $\overline{\ha}$ \\
        & [$10^6$] & [\emph{nObs}$>$1] & [mag] & [mag] & [mag] \\
        \midrule
        $r$, $i$, \ha & 159 & 91\% & 19.3 & 18.2 & 18.8  \\
        $r$, $i$ & 25 & 77\% & 21.2 & 19.9 & - \\
        $i$, \ha & 3 & 73\% & - & 19.6 & 20.2 \\
        $r$, \ha & 2 & 65\% & 20.6 & - & 20.1 \\
        $i$ & 15 & 43\% & - & 19.8 & - \\
        $r$ & 9 & 27\% & 21.1 & - & - \\
        \ha & 6 & 12\% & - & - & 19.9 \\
        \midrule
        Total  & 219 & 81\% & 19.6 & 18.6 & 18.9 \\
        \bottomrule
        \end{tabular}
\end{table}

% riha: 158831809 / 144339044
% ri: 24558323 / 18878201
% iha: 3422905 / 2495655
% rha: 1606867 / 1036607
% i: 15463660 / 6612351
% r: 9263422 / 2478774
% ha: 5844538 / 727732

@@

%Having explained how the catalogue was created,
We now offer an overview of the properties of the
catalogue by discussing  
(i) the recommended quality criteria,
(ii) the photometric uncertainties and repeatability,
and (iii) the source densities and the frequency of source blending.

@@

\section{Data access and source code}
\label{sec:dataaccess}

The catalogue is made available on our website (www.iphas.org)
as a collection of binary FITS tables,
each covering a $5\degr\times5\degr$ tile
of the footprint.
The catalogue can be queried in two ways.
First, will be made available through the
Vizier on-line service (http://vizier.u-strasbg.fr),
where it can be accessed using a web interface
and using Virtual Observatory (VO) protocols.
Second, on our website we provide a script
to ingest the binary FITS tables into a
{\sc PostgreSQL} database for users
who wish to mine the data locally.

The website also provides access to the pipeline-processed
imaging data, which we have updated to include
a new keyword, called {\sc photzp},
which contains the re-calibrated zeropoint.
The keyword can be used to convert the number counts $DN$,
which stored in the pixel values,
into Vega-based magnitudes $m$ using:
\begin{equation}
\begin{split}
   m  = \textsc{photzp} - 2.5 \log_{10}(DN).
\label{eqn:mag}
\end{split}
\end{equation}
That is, {\sc photzp} has been computed such that it
absorbs the required corrections for atmospheric extinction,
gain variations, exposure time, and the re-calibration.

A user will only have to apply a correction
for the flux lost outside

To convert the magnitude $m$ into fluxes,
we list in Table~\ref{tbl:flux} something.

% and the best-available
%astrometric solution in the image headers.

@@

Users of the data are reminded to cite both \citet{Drew2005}
and this paper.

Mean photon wavelength of the passband following eqn.~A14 in Bessel \& Murphy (2012):
\begin{equation}
\lambda_0 = \frac{\int{\lambda S(\lambda) d \lambda}}
{\int{S(\lambda)d\lambda}}
\end{equation}

Pivot wavelength of the passband following eqn.~A16 in Bessel \& Murphy (2012):
\begin{equation}
\lambda_p = \frac{\int{S(\lambda) \lambda d \lambda}}
{\int{\frac{S(\lambda)}{\lambda} d\lambda}}
\end{equation}

Mean monochromatic flux,
following eqn.~A11 in Bessel \& Murphy (2012):
\begin{equation}
f_\lambda = \frac{\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda}}
{\int{S(\lambda)\lambda d \lambda}}
\end{equation}
where $S(\lambda)$ includes the filter, atmosphere and QE.

\begin{table}
    \caption{Wavelength and mean monochromatic flux density for Vega in the IPHAS filter system.}
    \label{tbl:flux}
    \begin{tabular}{lrrrr}
    \toprule
     Band & $\lambda_0$ & $\lambda_p$ & EW  & $f_\lambda$   \\
     & \AA & \AA & \AA & erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$  \\
    \midrule
$r$ &  6223  &   6211  &  785.6  &  2.47e-9  \\
\ha &  6568  &   6568  &   59.6  &  1.81e-9  \\
$i$ &  7674  &   7674  &  759.9  &  1.30e-9   \\
    \bottomrule
    \end{tabular}
\end{table}

>>>>

Unlike the broadbands, 
standard-star photometry is not available in the literature 
for the \ha\ passband
and hence there is no formally recognised flux scale 
for it.
We can specify here, however, 
that the detected flux for Vega in the
IPHAS \ha\ filter is 3.14 magnitudes
less than the flux captured by the much broader
$r$ band which includes the \ha\ band.
Hence to assure zero colour relative to the broadbands,
we set the default zeropoint for the narrowband to be:
\begin{equation}
\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.14.
\label{eqn:zpha}
\end{equation}
For reference, Table~\ref{tbl:flux} details the flux of Vega
in the IPHAS filter system.
We refer to \citet{Drew2005} for a more comprehensive
discussion on the throughput curves of the bands.

\begin{table}
    \caption{Mean monochromatic flux of Vega in the IPHAS filter system,
             defined as 
             $\langle f_\lambda\rangle~=~\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda} / \int{S(\lambda)\lambda d \lambda}$,
             where $S(\lambda)$ is the photon response function 
             (which includes atmospheric transmission,
             filter transmission
             and CCD response)
             and $f_\lambda(\lambda)$ is the model SED for Vega
             by Kurucz (http://kurucz.harvard.edu/stars.html).
             For reference, we also provide the 
             filter equivalent width EW~=~$\int S(\lambda) d\lambda$,
             the mean photon wavelength 
             $\lambda_0~=~\int{S(\lambda)\lambda d\lambda} / \int{S(\lambda)d\lambda}$,
             and the pivot wavelength
             $\lambda_p~=~\sqrt{\int{S(\lambda)\lambda d\lambda} / \int{\frac{S(\lambda)}{\lambda} d\lambda}}$.
             These notations follow the definitions by \citet{Bessell2012}.
             After multiplying $\langle f_\lambda\rangle$ by the EW,
             we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
             than that received in $r$, 
          }
    \label{tbl:flux}
    \begin{center}
    \begin{tabular}{lccrc}
    \toprule
     Filter  & $\langle f_\lambda \rangle$  & \multicolumn{1}{c}{EW} &$\lambda_0$ & $\lambda_p$    \\
             & [erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$] & [\AA] & [\AA] & [\AA] \\
    \midrule
$r$ &  $2.47 \times 10^{-9}$  &  785.6 &  6223  &   6211   \\
\ha &  $1.81 \times 10^{-9}$ &   59.6 &  6568  &   6568    \\
$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7674   \\
    \bottomrule
    \end{tabular}
    \end{center}
\end{table}

@@

\begin{table}
    \caption{Breakdown of catalogue entries as a function of the band(s)
             in which the source was detected.
             We also show the fraction of `confirmed' objects,
             which we define as those sources detected in more than one field
             (usually the field pair partner).
             We also show the average magnitudes in each class.}
    \label{tbl:detections}
    \begin{center}
        \begin{tabular}{lrcccc}
        \toprule
        Band(s) & Sources & Confirmed & $\overline{r}$ & $\overline{i}$ & $\overline{\ha}$ \\
        & [$10^6$] & [\emph{nObs}$>$1] & [mag] & [mag] & [mag] \\
        \midrule
        $r$, $i$, \ha & 159 & 91\% & 19.3 & 18.2 & 18.8  \\
        $r$, $i$ & 25 & 77\% & 21.2 & 19.9 & - \\
        $i$, \ha & 3 & 73\% & - & 19.6 & 20.2 \\
        $r$, \ha & 2 & 65\% & 20.6 & - & 20.1 \\
        $i$ & 15 & 43\% & - & 19.8 & - \\
        $r$ & 9 & 27\% & 21.1 & - & - \\
        \ha & 6 & 12\% & - & - & 19.9 \\
        \midrule
        Total  & 219 & 81\% & 19.6 & 18.6 & 18.9 \\
        \bottomrule
        \end{tabular}
        % riha: 158831809 / 144339044
		% ri: 24558323 / 18878201
		% iha: 3422905 / 2495655
		% rha: 1606867 / 1036607
		% i: 15463660 / 6612351
		% r: 9263422 / 2478774
		% ha: 5844538 / 727732
	\end{center}
\end{table}

A total of 219 million primary detections satisfied
the above criteria and have been included in the catalogue.

Table~\ref{tbl:detections} details the number of bands
in which the catalogue entries are captured;
159 million sources are detected in all three filters (73 per cent),
30 million are detected in two filters (14 per cent),
and the remaining 30 million are single-band detections.
Table~\ref{tbl:detections} also presents the average magnitudes
and the fraction of `confirmed' objects,
which we define as those sources which have been detected in more than
one field
(recall that the majority of the survey area
is observed twice due to the field pair strategy).
We find that the single-band detections tend to show low confirmation rates,
which suggests that a significant fraction of these entries may be spurious detections.
Hence, we strongly recommend users not to rely on single-band objects
without inspecting the image data by eye.

Not all the single-band detections are spurious sources, however.
We note that the confirmation rate for $i$-band detections
is markedly better than for $r$ and \ha,
which is likely explained by the fact that $i$ is least
affected by interstellar extinction, and so the survey can occasionally pick up
highly-reddened objects in $i$ which are otherwise lost in $r$ and \ha.
Moreover, objects which are intrinsically very red may also be picked
up in $i$ alone, while faint objects with strong Balmer emission
may appear only in \ha.

@@

We now offer an overview of the properties of the
catalogue by discussing  
(i) the recommended quality criteria,
(ii) the photometric uncertainties and repeatability,
and (iii) the source densities and the frequency of source blending.

@@

\section{Catalogue and image access}
\label{sec:dataaccess}

The catalogue is made available through the
Vizier service (http://vizier.u-strasbg.fr),
where it can be queried using a web interface
and using Virtual Observatory (VO) protocols.
In addition, the catalogue can be downloaded
in its entirety from our website (www.iphas.org)
as a collection of binary FITS tables,
each covering a $5\degr\times5\degr$ tile
of the footprint.

We do not recommend using the catalogue
to study extended objects which
are larger than the
aperture diameters used in this work.
To enable the analysis of diffuse sources,
our website provides access to the pipeline-processed
imaging data. These have been updated to include
a new FITS header keyword, called {\sc photzp},
which contains the re-calibrated zeropoint.
This keyword can be used to convert the number counts $DN$,
i.e. the pixel values in the images,
into Vega-based magnitudes $m$ using:
\begin{equation}
\begin{split}
   m  = \textsc{photzp} - 2.5 \log_{10}(DN).
\label{eqn:mag}
\end{split}
\end{equation}
The {\sc photzp} value has been computed such that it
absorbs the required corrections for atmospheric extinction,
gain variations, exposure time, and the re-calibration shift.
Users will only need to apply an aperture correction, if necessary.

To estimate absolute narrow-band \ha\ fluxes from the image data,
we note that the integrated in-band energy flux for Vega 
in the IPHAS \ha\ filter 
is $1.52 \times 10^{-7}$~erg\,cm$^{-2}$\,s$^{-1}$ 
at the top of the Earth's atmosphere,
which is the flux obtained by folding the 
Vega model SED from Kurucz with the filter transmission curve 
(uncorrected for the atmosphere and detector quantum efficiency,
which would otherwise scale down the narrowband flux by 0.707).
This implies that the in-band flux corresponding to
zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
when the \ha\ magnitude for Vega is set by convention to 0.03 \citep{Fukugita1996}.

The image repository on our website
is accompanied by an archive of detection tables
-- one for each image --
which may be used to retrieve secondary detections that are not included
in the DR2 source catalogue.
We warn that this set of original images and detection tables
includes data that did not pass quality control,
and has not been globally re-calibrated.
Such data are flagged and must be used with great caution.

@@
